Combining information from photometric and spectroscopic data:\\ propogating uncertainties in spectrophotometric calibration

### 2015-04-29 tex/ms.tex
updates to bib, small changes throughout text, add some plotting code

If you don't believe the reported data noise variances $\sigma_n^2$,
it often makes sense to add in quadrature a ``jitter'' variance
$s^2$.  For photometric data this ``jitter'' is often best described
as a fractional error, $(\hat{s}\mu(\theta))^2$, corresponding for example
to an additional magnitude zero-point uncertainty.
This is a simple modification of the chi-squared log-likelihood:
\begin{eqnarray}\label{eq:photometricLF}
\ln p(d\given\theta,s^2) &=& -\frac{1}{2}\,\sum_{n=1}^N
                             \left[\frac{[d_n - \mu_n(\theta)]^2}
                             {[\sigma_n^2 + s^2]} + 
                             \ln(2\pi\,[\sigma_n^2 + s^2]) \right]
\quad .
\end{eqnarray}
This modification shows the necessity of including the log-variance
term.
This---equation~(\ref{eq:photometricLF})---is in fact the
log-likelihood function we are going to use for the photometric
measurements in what follows.  In this formulation, if $\theta$ are
the parameters of interest, the jitter variance $\hat{s}^2\equiv
s/\mu(\theta)$ is a nuisance noise-model parameter representing
additonal fractional uncertainty in the flux.  The Gaussian Process
will be a generalization of this modification.

@@

@@

Next we will \emph{not} assume that the calibration is perfectly known
and we will therefore allow the calibration parameters $\alpha$ in our
model to vary.  In Figure
\ref{fig:speconly_calibrated} we show the result of our inference from
the spectrum and a single photometric measurement (SDSS $g$) for a
perfectly calibrated mock spectrum. To demonstrate the quality of the
fit we show both the mock input spectrum and predictions of the
observed spectra that result from several draws of the model
parameters $[\theta, \alpha]$ from posterior distribution.  We also
show samples of the inferred calibration vector,
\begin{eqnarray} 
\frac{d}{\mu(\theta)} & = & e^{f(\alpha) + \tilde{\Delta}}
\quad ,
\end{eqnarray}
where $\tilde{\Delta}$ is the Gaussian Process mean prediction for the
logarithmic residual between the data and the model.  For the
perfectly calibrated spectrum the true calibration vector (shown in
black) is one everywhere.  The spread of the inferred calibration
vectors approximately indicates our uncertainty in the spectrophotometric
calibration when no additional calibration information is available
and only a single photometric point is available.

@@

\subsection{Real Data: NGCXXXX} 
We first infer the parameters of interest $\theta$ for a calibrated
spectrum of one of the clusters, \excluster, using the full YY-band
photometry. In Figure \ref{fig:inferred_spectrum} we show the spectra
that result from draws of the parameters $[\theta,\alpha]$ from the
posterior distribution, as well as several of the individual
components of the model constructed from each draw. Figure
\ref{fig:inferred_spectrum} also shows the calibrated spectrum on
which the priors were conditioned, and residuals between the posterior
models and the observed data.

@@

It is possible to reconstruct the inferred total multiplicative
calibration vector for each draw of the model parameters.  Rearranging
equation \ref{eq:spectroscopicLF} yields
\begin{eqnarray} 
\frac{d}{\mu(\theta)} & = & e^{f(\alpha) + \tilde{\Delta}}
\quad ,
\end{eqnarray}
for the total effective multiplicative calibration, where
$\tilde{\Delta}$ is now the Gaussian Process mean prediction for the
logarithmic residual between the data and the model. The inverse of
this ratio is shown in the left panel of Figure
\ref{fig:inferred_calib} as our inferred total multiplicative
calibration function; if the spectrophotometric calibration were
perfect and we had correctly inferred it, then we would expect this
vector to be constant as a function of wavelength.

@@

We have demonstrated, using using both mock and real spectra of
galactic stellar clusters, that we can combine photometric and
spectroscopic data in a principled way to infer physical properties.
We do not resort to an arbitrary down-weighting or up-weighting of
certain data. Instead, we explicitly model our uncertain knowledge of
the spectral calibration and infer this calibration jointly with the
physical properties. For a number of applications this combination of
spectroscopy and photometry reduces the need for spectrophotometric
calibration entirely, such that the less expensive photometric
calibration is sufficient.  Additional calibration information, if it
exists, can be incorporated into informative priors on the calibration
parameters of the model, inform the construction of the calibration
model (including the form of the covariance kernel), or approximately
remove gross features that are difficult to model parametrically or
via a Gaussian Process.

@@

Our strategy is easily extensible to more complicated models, which
include for example night sky-line emission in the spectroscopy (or
residuals from an intial sky-line subtraction) and uncertainties in
wavelength calibration \citep[e.g.,][]{walker15}. Furthermore, more complex physical models may
be considered for the mean object model, for example including SFH
parameters so as to model galaxies.  {\color{blue} DISCUSS MORE
COMPLICATED COVARIANCE KERNELS, MULTIPLE KERNELS, etc.  MODEL
CHECKING.}

@@

Limitations.  The strategy we have advocated has several limitations.
First, it requires a generative model for the spectrum of the object
of interest, and is thus ill-suited to discovery projects.  Second,
care must be taken to avoid absorbing interesting model deficiencies
into the calibration model. For example, in the model we have
described, it is possible that broad shallow features (for example due
to molecular bands or absorption line complexes) will be absorbed into
the Gaussian process. Third, we have made the assumption that the
spectroscopy and the photometry are sampling the same populations.  In
practice, differences between spectroscopic and photometric apertures
often mean that the photometry and the spectroscopy are sampling
slightly different physical regions.  The correct response to this
situation is to explicitly model the differences, incorporating prior
information about the spatial variation of physical properties.
Fourth, the method will break down at very low S/N due to the
logarithmic formulation of the residual in
equation~(\ref{eq:spectroscopicLF}).  Finally, this method is
computationally intensive and expensive, since we are constantly
inverting largish matrices.

@@

>>>>

If you don't believe the reported data noise variances $\sigma_n^2$,
it often makes sense to add in quadrature a ``jitter'' variance $s^2$.
For photometric data this ``jitter'' is often best described as a
fractional error, $s^2 \equiv \hat{s}^2\mu(\theta)^2$, corresponding
for example to an additional magnitude zero-point uncertainty.  This
is a simple modification of the chi-squared log-likelihood:
\begin{eqnarray}\label{eq:photometricLF}
\ln p(d\given\theta,s^2) &=& -\frac{1}{2}\,\sum_{n=1}^N
                             \left[\frac{[d_n - \mu_n(\theta)]^2}
                             {[\sigma_n^2 + s^2]} + 
                             \ln(2\pi\,[\sigma_n^2 + s^2]) \right]
\quad .
\end{eqnarray}
This modification shows the necessity of including the log-variance
term.
This---equation~(\ref{eq:photometricLF})---is in fact the
log-likelihood function we are going to use for the photometric
measurements in what follows.  In this formulation, if $\theta$ are
the parameters of interest, the jitter variance $\hat{s} = 
s/\mu(\theta)$ is a nuisance noise-model parameter representing
additonal fractional uncertainty in the flux.  The Gaussian Process
will be a generalization of this modification.

@@

{\color{blue} BDJ This paragraph should explain why we are using the
GP, not just how it works.}
The multiplicative calibration vector can be explicitly modeled, for
example using polynomials.  However, for complex calibration vectors
over large wavelength ranges, the number of parameters required to
model the calibration can become quite large (e.g., very high order
polynomials) and the correct parameterization may be unknown.  It is
therefore desirable to introduce a non-parameteric model of the
calibration.  We do this using a Gaussian Process: prior over
functions, analytically marginalizable, covariant fractional
uncertainties the same as multiplying by a function (covariant
uncertainties the same as adding a function)

@@

Next we will \emph{not} assume that the calibration is perfectly known
and we will therefore allow the calibration parameters $\alpha$ in our
model to vary.  In Figure
\ref{fig:speconly_calibrated} we show the result of our inference from
the spectrum and a single photometric measurement (SDSS $g$) for a
perfectly calibrated mock spectrum. To demonstrate the quality of the
fit we show both the mock input spectrum and predictions of the
observed spectra that result from several draws of the model
parameters $[\theta, \alpha]$ from posterior distribution.  We also
show samples of the inferred calibration vector,
\begin{eqnarray} \label{eq:calibration}
\frac{d}{\mu(\theta)} & = & e^{f(\alpha) + \tilde{\Delta}}
\quad ,
\end{eqnarray}
where $\tilde{\Delta}$ is the Gaussian Process mean prediction for the
logarithmic residual between the data and the model \citep{RW}.  For the
perfectly calibrated spectrum the true calibration vector (shown in
black) is one everywhere.  The spread of the inferred calibration
vectors approximately indicates our uncertainty in the spectrophotometric
calibration when no additional calibration information is available
and only a single photometric point is available.

@@

\subsection{Real Data: \excluster} 
We first infer the parameters of interest $\theta$ for a calibrated
spectrum of one of the clusters, \excluster, using the full YY-band
photometry. In Figure \ref{fig:inferred_spectrum} we show the spectra
that result from draws of the parameters $[\theta,\alpha]$ from the
posterior distribution, as well as several of the individual
components of the model constructed from each draw. Figure
\ref{fig:inferred_spectrum} also shows the calibrated spectrum on
which the priors were conditioned, and residuals between the posterior
models and the observed data.

@@

It is possible to reconstruct the inferred total multiplicative
calibration vector for each draw of the model parameters (Equation
\ref{eq:calibration}) If the spectrophotometric calibration were
perfect and we had correctly inferred it, then we would expect this
vector to be constant as a function of wavelength.

@@

We have demonstrated, using using both mock and real spectra of
galactic stellar clusters, that we can combine photometric and
spectroscopic data in a principled way to infer physical properties.
We do not resort to an arbitrary down-weighting or up-weighting of
certain data. Instead, we explicitly model our uncertain knowledge of
the spectral calibration and infer this calibration jointly with the
physical properties. For a number of applications this combination of
spectroscopy and photometry reduces the need for spectrophotometric
calibration entirely, such that the less expensive photometric
calibration is sufficient.  Additional calibration information, if it
exists, can be incorporated into informative priors on the calibration
parameters of the model, inform the construction of the calibration
model (including the form of the covariance kernel), or approximately
remove gross features that are difficult to model parametrically or
via a Gaussian Process.  {\color{blue} Discuss more complicated
covariance kernels, multiple kernels}

@@

Our strategy is easily extensible to more complicated models, which
include for example night sky-line emission in the spectroscopy
\citep[e.g.,][]{gullikson14} or residuals from an intial sky-line
subtraction and uncertainties in wavelength calibration
\citep[e.g.,][]{walker15}. Furthermore, more complex physical models
may be considered for the mean object model, for example including SFH
parameters so as to model galaxies. {\color{blue} Discuss background
modeling}

@@

Limitations.  The strategy we have advocated has several limitations.
First, it requires a generative model for the spectrum of the object
of interest, and is thus ill-suited to discovery projects.  Second,
care must be taken to avoid absorbing interesting model deficiencies
into the calibration model. For example, in the model we have
described, it is possible that broad shallow features (for example due
to molecular bands or absorption line complexes) will be absorbed into
the Gaussian process if not properly included in the mean
model. Third, we have made the assumption that the spectroscopy and
the photometry are sampling the same populations.  In practice,
differences between spectroscopic and photometric apertures often mean
that the photometry and the spectroscopy are sampling slightly
different physical regions.  The correct response to this situation is
to explicitly model the differences, incorporating prior information
about the spatial variation of physical properties.  Fourth, the
method will break down at very low S/N due to the logarithmic
formulation of the residual in equation~(\ref{eq:spectroscopicLF}).
Finally, this method is computationally intensive and expensive, since
we are constantly inverting largish matrices.

@@

\bibitem[Gullikson et al.(2014)]{gullikson14} Gullikson, K., 
Dodson-Robinson, S., \& Kraus, A.\ 2014, \aj, 148, 53
